\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[LY1]{fontenc}
\usepackage{berenis}
\usepackage{amsmath, amssymb}
\usepackage{anyfontsize}
\title{Ezekiel}
\author{Svadrut Kukunooru \& Danny Chung}


\begin{document}
    \maketitle
    The prophetic model is split into 4 parts:
    \begin{itemize}
        \item \textit{Revelation}: How the prophet originally gets the prophecy from God
        \item \textit{Prophecy}: How the prophet interprets the revelation and what he tells other people
        \item \textit{Supernatural Confirmation}: Making the prophecy come true
        \item \textit{Feedback}: Feedback from both God and the people.
    \end{itemize}
\section{Revelation}%
\label{sec:Revelation}
Several visions that Ezekiel has of God are detailed throughout the \textit{Book of Ezekiel}. For his first vision, God approaches Ezekiel as a warrior riding a chariot drawn by creatures with four faces and four wings. He says to Ezekiel, "Son of man, I am sending you to the Israelites."(2:3), making Ezekiel a prophet and a "watchman" of Israel. Ezekiel's relevations follow a parabolic trajectory in mood. First, Ezekiel sees God enthroned in the Temple, but then leave because pagan worship is being performed inside the Temple. Next there are metaphors of Israel being a prostitute and how Israel is dying. However, Ezekiel sees Israel rise from the dead in "the valley of dry bones" and Israel's enemies destroyed. Finally, Ezekiel sees God enthroned in the "Third" Temple. 
\section{Prophecy}%
\label{sec:Prophecy}
Ezekiel first spent five years prophesying the destruction of Jerusalem, much to the ridicule of the majority of the population. Next, he prophesies that Israel will beat its foreign enemies; in this case, the nation-tribes of Gog and Magog. Next, he basically prophesies the comeback of Israel and the building of a "Third Temple" for the Jews. 
\section{Supernatural Confirmation}%
\label{sec:Supernatural Confirmation}
In terms of all the prophecies that Ezekiel was given, Babylonians came to destroy Jerusalem five years after Ezekiel began to prophesy about such an event. Once that occurred, he exiled himself to Babylon and was a known prophet through Israel. After the nation-tribes of Gog and Magog were destroyed following his prophecy, Ezekiel brought Israel back together following his two prophecies that the third temple would be built and through ending the Jewish exile, a new city and temple were built. 
\section{Feedback}%
\label{sec:Feedback}
For the first of Ezekiel's prophecies, he is given much ridicule by the people of Jerusalem for thinking that Jerusalem might fall. Unfortunately, he's right -- after a lengthy siege, Jerusalem is taken by the Babylonians. Ezekiel is known, more than other prophets, to undertake symbolic actions and have strange visions/trances; in this way, he was quite similar to Ramakrishna of Dineshwar. As a result, most people saw his views as irrelevant until they were proven true. Later, they began to trust in this new covenant to God. 
\end{document}
